---
title: | 
 | Patient Molecular Profiling Caris Data:  
 | R/Shiny DataViz Showcase 
author: "Yi Chen"
date:    '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
runtime: shiny
---

```{r echo=FALSE,warning=FALSE,cache=FALSE,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DT)
library(markdown)
library(shinythemes)
library(corrplot)
source("Circle plot.R")
library(dplyr)

```


### Introduction

**This Shiny App allows for the following displays:**

1. **Heatmaps/tables of biomarkers for _disease_ and _disease category_, by _technology_ and _actionability_:** 
    * Heatmap
        + Sort the top 50 biomarkers that have been most mutated among patients
        + Color can be Red/Green or Blue/Yellow
    * Table
        + Table has first 12 columns from raw data
        + Table allows for sorting and searching
  
2. **Biomarker correlation circle plot by _technology_:**
    * Only consider predetermined set of biomarkers (may expand in the future) filtered by technology

### Shiny App
```{r echo=FALSE,warning=FALSE,cache=FALSE}
shinyApp(
  options=list(width = "110%",height=2300),
  
# Heatmap by Disease Categories
ui = navbarPage("Theme",
            theme = shinytheme("cerulean"),
            tabPanel("Disease Category",
            titlePanel("Heatmap and Table by Disease Categories"),
            fluidRow(column(3,
                             h4("Disease Category Explorer"),
                             # Copy the chunk below to make a group of checkboxes
                    checkboxGroupInput("checkdisease.dc", "Disease Categories",
                                       choices = unique(my.molecular[!is.na(my.molecular$Disease.Category),]$Disease.Category)),
                    radioButtons("Color.dc", "Color Type",choices=list("Red Green"="red_green","Blue Yellow"="blue_lightgoldenrod1"))

                             ),
                    column(5,offset=2,selectInput("Tech.dc", "Technology Type",choices = sort(unique(my.molecular$Technology[(!is.na(my.molecular$Disease.Category))&(!is.na(my.molecular$Disease))]))),
                    checkboxGroupInput("actionability.dc", "Actionability Assessment Levels",
                                       choices = unique(Actionability$Actionability)),

                    helpText("Level 1: can be targeted by approved drugs
"),
helpText("Level 2: can be targeted by investigational drugs
"))),
tweaks <-
                    list(tags$head(tags$style(HTML("
                                 .multicol {
                                   height: 63px;
                                   -webkit-column-count: 8; /* Chrome, Safari, Opera */
                                   -moz-column-count: 8;    /* Firefox */
                                   column-count: 8;
                                   -moz-column-fill: auto;
                                   -column-fill: auto;
                                 }
                                 "))
                                 )) ,
fluidRow(column(12,

                           controls <-list(tags$div(align = 'left',
                class = 'multicol',
                checkboxGroupInput(inputId  = 'functional.dc',
                label    = "Functional Category",
                choices  = sort(unique(Functional$Category)),
                inline   = FALSE))))),


                    tags$hr(),
                     fluidRow(
      tabsetPanel(type = "tabs",
        tabPanel("Heatmap", 
                 plotOutput("heatmap.dc",height="600px",width="100%"),
                 h5("Reference"),verbatimTextOutput("value.dc")),
        tabPanel("Table", DT::dataTableOutput('ex1'))
      ))),


           tabPanel("Disease",
                    titlePanel("Heatmap and Table by Disease"),
                    # Copy the chunk below to make a group of checkboxes
                    tweaks2 <-
                      list(tags$head(tags$style(HTML("
                                                     .multicol2 {
                                                     height: 450px;
                                                     -webkit-column-count: 4; /* Chrome, Safari, Opera */
                                                     -moz-column-count: 4;    /* Firefox */
                                                     column-count: 4;
                                                     -moz-column-fill: auto;
                                                     -column-fill: auto;
                                                     }
                                                     "))
                      )),
#
                    controls2 <-
                      list(h4("Disease Explorer"),
                           tags$div(align = 'left',
                                    class = 'multicol2',
                                    checkboxGroupInput(inputId  = 'checkdisease.d',
                                                       label    = "Disease",
                                                       choices  = sort(unique(my.molecular[!is.na(my.molecular$Disease),]$Disease)),
                                                       inline   = FALSE))),
                    fluidRow(
                      column(3,
                             radioButtons("Color.d", "Color Type",choices=list("Red Green"="red_green",
                                                                      "Blue Yellow"="blue_lightgoldenrod1")),selectInput("Tech.d", "Technology Type",choices = sort(unique(my.molecular$Technology)))
                             ),
                      column(6,offset=3,

                    checkboxGroupInput("actionability.d", "Actionability Assessment Levels",
                                       choices = unique(Actionability$Actionability)),

                    helpText("Level 1: can be targeted by approved drugs
"),
helpText("Level 2: can be targeted by investigational drugs
")
                             )
                    ),
tweaks <-
                    list(tags$head(tags$style(HTML("
                                 .multicol {
                                   height: 63px;
                                   -webkit-column-count: 8; /* Chrome, Safari, Opera */
                                   -moz-column-count: 8;    /* Firefox */
                                   column-count: 8;
                                   -moz-column-fill: auto;
                                   -column-fill: auto;
                                 }
                                 "))
                                 )) ,
fluidRow(column(12,

                controls <-list(tags$div(align = 'left',
                class = 'multicol',
                checkboxGroupInput(inputId  = 'functional.d',
                label    = "Functional Category",
                choices  = sort(unique(Functional$Category)),
                inline   = FALSE))))),
                    tags$hr(),

                    fluidRow(
                      plotOutput("heatmap.d",height="600px",width="100%"),
                      h5("Reference"),verbatimTextOutput("value.d"),
                      DT::dataTableOutput('ex2')

                    )),
           tabPanel("Correlation Plot by Technology",
                    titlePanel("Biomarker Correlation Circle Plot by Technology"),
                    # Copy the chunk below to make a group of checkboxes

                    
                        fluidRow(column(3,wellPanel(
                          selectInput("Tech.cir", "Technology Type",choices = name),
                    selectInput("Tech.cir2", "Technology Type Comparing",choices = name),


                    helpText("Red for positive and Blue for negative correlation"),
                    helpText("The higher correlation, the darker the color"),
                    helpText("Circle plot color based on quantile for both positive and negative values")
                    
                    # helpText("Color classification is based on quantile: positive: (0.132,0.280,0.348,0.576), negative: (-0.0024,-0.0070,-0.0108,-0.016)")
                        )),
                    


                    # tags$hr(),

                    mainPanel(
        tabsetPanel(type = "tabs",
        tabPanel("Circle Plot", plotOutput("circle.dc",height="800",width="100%")),
        tabPanel("Correlation Matrix", plotOutput("corplot.dc",height="800",width="100%"))
      ))))),




## Update filtered dataset and calculation of correlation

server=function(input, output){
  output$heatmap.dc <- renderPlot ({
    
    label_change<-c( as.character(Legend.label$Positive[Legend.label$Label==input$Tech.dc]),as.character(Legend.label$Negative[Legend.label$Label==input$Tech.dc]))
    data<-my.molecular[(!is.na(my.molecular$Disease.Category))&(!is.na(my.molecular$PatientID)),]
    update.data<-data[which(data$Technology%in%input$Tech.dc),]
    gene<-unique(update.data$Biomarker)
    # Heatmap by Disease Categories
    
    if(is.null(input$checkdisease.dc)){
      
      if(is.null(input$actionability.dc)){
      if (is.null(input$functional.dc)){
        if (length(gene)<50){
        plot.heatmap(update.data,
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease.Category',facetangle=0,legends=label_change,xlab = 'Patient', ylab = 'Gene',color=input$Color.dc,
                     title='')
        
        
      }
      
      else {
        sum.gene<-sapply(gene,function(x) sum(update.data[update.data$Biomarker==x,]$Positive))
        names(sum.gene)<-gene
        select.gene<-names(sort(sum.gene,decreasing = TRUE))[1:50]
        
          plot.heatmap(update.data[which(update.data$Biomarker%in%select.gene),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',facetangle=0,
                     category.x = 'Disease.Category',color=input$Color.dc,
                     xlab = 'Patient', ylab = 'Gene',legends=label_change,
                     title='')
          }
        }
      else{
        if (!(nrow(update.data[which(update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.dc]),]) > 0)){
    warning('There is no biomarker in this Functional Category')
      }
      else{
        
      # if (length(gene)<50){
        plot.heatmap(update.data[which(update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.dc]),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease.Category',facetangle=0,legends=label_change,xlab = 'Patient', ylab = 'Gene',color=input$Color.dc,
                     title='')
      
        
      }}}
    else{
      action.gene<-unique(Actionability$Gene[Actionability$Actionability%in%input$actionability.dc])
      if (!(nrow(update.data[which(update.data$Biomarker%in%action.gene),]) > 0)){
    warning('There is no biomarker in this level.')
      }
      else{
        if (is.null(input$functional.dc)){
          plot.heatmap(update.data[which(update.data$Biomarker%in%action.gene),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease.Category',facetangle=0,legends=label_change,
                     xlab = 'Patient', ylab = 'Gene',color=input$Color.dc,
                     title='')
        }
        else{
          if (!(nrow(update.data[which(update.data$Biomarker%in%action.gene&update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.dc]),]) > 0)){
    warning('There is no biomarker in this Functional Category')
      }
      else{
        plot.heatmap(update.data[which(update.data$Biomarker%in%action.gene& update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.dc]),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease.Category',facetangle=0,legends=label_change,
                     xlab = 'Patient', ylab = 'Gene',color=input$Color.dc,
                     title='')
      }
        }
      }
      
    }}
    
    else{
      
      update.data<-update.data[which(update.data$Disease.Category%in%input$checkdisease.dc),]
      gene<-unique(update.data$Biomarker)

      if(is.null(input$actionability.dc)){
              if (is.null(input$functional.dc)){
                if (length(gene)<50){
        
        
        plot.heatmap(update.data,legends=label_change,
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',facetangle=0,
                     category.x = 'Disease.Category',color=input$Color.dc,
                     xlab = 'Patient', ylab = 'Gene',
                     title='')
        
        }
      else {
        sum.gene<-sapply(gene,function(x) sum(update.data[which( update.data$Biomarker%in%x),]$Positive))
        names(sum.gene)<-gene
        select.gene<-names(sort(sum.gene,decreasing = TRUE))[1:50]
        
        plot.heatmap(update.data[which(update.data$Biomarker%in%select.gene),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',facetangle=0,
                     category.x = 'Disease.Category',color=input$Color.dc,
                     xlab = 'Patient', ylab = 'Gene',legends=label_change,
                     title='')
      }
              }
        
    else{
      if (!(nrow(update.data[which(update.data$Biomarker%in%Functional$Gene[Functional$Category==input$functional.dc]),]) > 0)){
    warning('There is no biomarker in this Functional Category')
      }
      else{
        plot.heatmap(update.data[which(update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.dc]),],legends=label_change,           x = 'PatientID', y = 'Biomarker', response = 'Positive',facetangle=0,
                     category.x = 'Disease.Category',color=input$Color.dc,
                     xlab = 'Patient', ylab = 'Gene',
                     title='')
    }}}
      
      
      else{
        
        action.gene<-unique(Actionability$Gene[Actionability$Actionability%in%input$actionability.dc])
        if (!(nrow(update.data[which(update.data$Biomarker%in%action.gene),]) > 0)){
            warning('There is no biomarker in this level.')
        }
        else{
          if (is.null(input$functional.dc)){
            plot.heatmap(update.data[which(update.data$Biomarker%in%action.gene),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease.Category',facetangle=0,legends=label_change,
                     xlab = 'Patient', ylab = 'Gene',color=input$Color.dc,
                     title='')
          }
          else{
          
          if (!(nrow(update.data[which(update.data$Biomarker%in%action.gene&update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.dc]),]) > 0)){
    warning('There is no biomarker in this Functional Category')
      }
      else{
          
        plot.heatmap(update.data[which(update.data$Biomarker%in%action.gene& update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.dc]),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease.Category',facetangle=0,legends=label_change,
                     xlab = 'Patient', ylab = 'Gene',color=input$Color.dc,
                     title='')
          }
      
      }}
    
  }}}) 
  
  output$ex1<-DT::renderDataTable( 
    

        if(is.null(input$checkdisease.dc)){
          

      DT::datatable(my.molecular[(!is.na(my.molecular$Disease.Category))&my.molecular$Technology==input$Tech.dc,1:12],options=list(pageLength=10))
    }
    else{

      DT::datatable(my.molecular[(!is.na(my.molecular$Disease.Category))&(my.molecular$Disease.Category==input$checkdisease.dc)&(my.molecular$Technology==input$Tech.dc),1:12],options=list(pageLength=10))
    }
  )
  
    
  output$value.dc<-renderText({

        
    paste(as.character(Legend.label$Note[as.character(Legend.label$Label)==input$Tech.dc]),
    "White color in heatmap means missing value",sep="\n")
  })
 
  output$circle.dc<-renderPlot( {
    name<-sort(unlist(unname(sapply(names(correlation),function(x) x[which(sum(!is.na(correlation[[x]]))!=0)]))))
    color.division<-unique(unlist(correlation))
    color.pos<-color.division[color.division>0]
    color.neg<-color.division[color.division<0]
    pos<-quantile(color.pos,na.rm=TRUE)
    neg<-quantile(color.neg,na.rm=TRUE)



    circle.plot2(input$Tech.cir,input$Tech.cir2)
    
  }, height = 780, width = 750)
  
  
  output$corplot.dc<-renderPlot({
    par(mar=c(5.1,4.1,4.1,2.1))
    

    # name<-sort(unlist(unname(sapply(names(correlation),function(x) x[which(sum(!is.na(correlation[[x]]))!=0)]))))
    # color.division<-unique(unlist(correlation))
    # color.pos<-color.division[color.division>0]
    # color.neg<-color.division[color.division<0]
    # pos<-quantile(color.pos,na.rm=TRUE)
    # neg<-quantile(color.neg,na.rm=TRUE)


    
    
    plot.data<-plot.data.pre(input$Tech.cir,input$Tech.cir2)
  
    if (!is.matrix(plot.data)){
  warning(plot.data)
}
else{
  if (input$Tech.cir==input$Tech.cir2){
        diag(plot.data)<-1

  }
      colnames(plot.data)<-paste(colnames(plot.data),input$Tech.cir,sep=" ")
      rownames(plot.data)<-paste(rownames(plot.data),input$Tech.cir2,sep=" ")
      corrplot(plot.data,tl.col="black",tl.cex=1,insig = "blank",method="color",na.label = "NA",col=rev(brewer.pal(8,"RdYlBu")))}
    }, height = 780, width = 750)
  
  output$heatmap.d <- renderPlot ({

    
    # Heatmap by Disease
    
    label_change<-c(as.character(Legend.label$Positive[Legend.label$Label==input$Tech.d]),as.character(Legend.label$Negative[Legend.label$Label==input$Tech.d]))
    data<-my.molecular[which((!is.na(my.molecular$Disease))&(!is.na(my.molecular$PatientID))),]
    update.data<-data[which(data$Technology==input$Tech.d),]
    gene<-unique(update.data$Biomarker)
    
    # Heatmap by Disease
    if(is.null(input$checkdisease.d)){
      if(is.null(input$actionability.d)){
        if (is.null(input$functional.d)){
          if (length(gene)<50){
        
        
        plot.heatmap(update.data,legends=label_change,
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease',color=input$Color.d,
                     xlab = 'Patient', ylab = 'Gene',
                     title='')
        
        
      }
      else {
        sum.gene<-sapply(gene,function(x) sum(update.data[update.data$Biomarker==x,]$Positive))
        names(sum.gene)<-gene
        select.gene<-names(sort(sum.gene,decreasing = TRUE))[1:50]
        
        plot.heatmap(update.data[which(update.data$Biomarker%in%select.gene),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease',color=input$Color.d,
                     xlab = 'Patient', ylab = 'Gene',legends=label_change,
                     title='')
        
        
        
      }}
        else{
        if (!(nrow(update.data[which(update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.d]),]) > 0)){
    warning('There is no biomarker in this Functional Category')
      }
      else{
      plot.heatmap(update.data[which(update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.d]),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease',color=input$Color.d,
                     xlab = 'Patient', ylab = 'Gene',legends=label_change,
                     title='')
        }}
        }

      
    else{
      action.gene<-unique(Actionability$Gene[Actionability$Actionability%in%input$actionability.d])
      if (!(nrow(update.data[which(update.data$Biomarker%in%action.gene),]) > 0)){
    warning('There is no biomarker in this level.')
      }
      else{
        if (is.null(input$functional.d)){
        plot.heatmap(update.data[which(update.data$Biomarker%in%action.gene),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease',legends=label_change,
                     xlab = 'Patient', ylab = 'Gene',color=input$Color.d,
                     title='')
        }
        
        else{
          if (!(nrow(update.data[which(update.data$Biomarker%in%action.gene&update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.d]),]) > 0)){
    warning('There is no biomarker in this Functional Category')
      }
      else{
        plot.heatmap(update.data[which(update.data$Biomarker%in%action.gene& update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.d]),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease',legends=label_change,
                     xlab = 'Patient', ylab = 'Gene',color=input$Color.d,
                     title='')
        

        }
    }}}}
    
    else{
      if (!(nrow(update.data[which(update.data$Disease%in%input$checkdisease.d),]) > 0)){
    warning('There is no patient in the sample for this Technology.')
      }
      else {
        update.data<-update.data[which(update.data$Disease%in%input$checkdisease.d),]
        gene<-unique(update.data$Biomarker)
      if(is.null(input$actionability.d)){
        if (is.null(input$functional.d)){
          if (length(gene)<50){
        plot.heatmap(update.data,
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease',color=input$Color.d,
                     xlab = 'Patient', ylab = 'Gene',legends=label_change,
                     title='')
        
        }
        else{
          sum.gene<-sapply(gene,function(x) sum(update.data[which(update.data$Biomarker%in%x),]$Positive))
        names(sum.gene)<-gene
        select.gene<-names(sort(sum.gene,decreasing = TRUE))[1:50]
        
        
        plot.heatmap(update.data[which(update.data$Biomarker%in%select.gene),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease',color=input$Color.d,
                     xlab = 'Patient', ylab = 'Gene',legends=label_change,
                     title='')
        
        
        }
          
        }
        else{
          if (!(nrow(update.data[which(update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.d]),]) > 0)){
    warning('There is no biomarker in this Functional Category')
      }
      else{
      
        plot.heatmap(update.data[which(update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.d]),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease',legends=label_change,
                     xlab = 'Patient', ylab = 'Gene',color=input$Color.d,
                     title='')
        }
        
      
      }}
      else{
        
        action.gene<-unique(Actionability$Gene[Actionability$Actionability%in%input$actionability.d])
        if (!(nrow(update.data[which(update.data$Biomarker%in%action.gene),]) > 0)){
            warning('There is no biomarker in this level.')
        }
      else {
        if (is.null(input$functional.dc)){
            plot.heatmap(update.data[which(update.data$Biomarker%in%action.gene),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease',color=input$Color.d,
                     xlab = 'Patient',ylab = 'Gene',legends=label_change,
                     title='')
          
        }
        else{
      if (!(nrow(update.data[which(update.data$Biomarker%in%action.gene&update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.d]),]) > 0)){
    warning('There is no patient in the sample for this Technology.')
      }
      else {
        plot.heatmap(update.data[which(update.data$Biomarker%in%action.gene & update.data$Biomarker%in%Functional$Gene[Functional$Category%in%input$functional.d]),],
                     x = 'PatientID', y = 'Biomarker', response = 'Positive',
                     category.x = 'Disease',color=input$Color.d,
                     xlab = 'Patient',ylab = 'Gene',legends=label_change,
                     title='')
        
      
        }
      
    }}}}}
  })
  output$value.d<-renderText({
    paste(as.character(Legend.label$Note[as.character(Legend.label$Label)==input$Tech.d]),
    "White color in heatmap means missing value",sep="\n")
  })
  
  output$ex2<-DT::renderDataTable(
    if(is.null(input$checkdisease.d)){

      DT::datatable(my.molecular[which(my.molecular$Technology==input$Tech.d),1:12],options=list(pageLength=10))
    }
    else{

      DT::datatable(my.molecular[which(my.molecular$Disease==input$checkdisease.d & my.molecular$Technology==input$Tech.d),1:12],options=list(pageLength=10))
    }
    
  )

})
```

